---
title: "section"
output: html_document
date: "2026-04-30"
---

---

title: "Startup Success Analyzer"
output: html_document
date: "2026-04-30"
------------------

# Startup Success Analyzer

---

## 1. Problem Statement

The objective of this project is to analyze startup success using statistical techniques.
We aim to understand whether startup sectors differ in terms of growth and identify patterns using multivariate analysis.

---

## 2. Data Cleaning

```{r}
startup_funding <- read.csv("C:/Users/aagma_r95jbd4/Downloads/startup_funding.csv")

colnames(startup_funding) <- c("Company", "Sector", "Funding", "Investor", "Year", "City")

startup_funding$Sector <- as.factor(startup_funding$Sector)

startup_funding <- na.omit(startup_funding)
```

---

## 3. Feature Engineering

```{r}
set.seed(123)

startup_funding$Revenue <- startup_funding$Funding * runif(nrow(startup_funding), 0.5, 2)

startup_funding$Employee <- round(startup_funding$Funding / 1000000 * runif(nrow(startup_funding), 5, 20))

startup_funding$Marketing <- startup_funding$Funding * runif(nrow(startup_funding), 0.1, 0.3)

startup_funding$Operational_Years <- 2024 - startup_funding$Year
```

**Explanation:**
Additional variables were created to simulate realistic startup metrics such as revenue, employee count, and marketing spend.

---

## 4. Exploratory Data Analysis (EDA)

### Funding Distribution Across Sectors

```{r}
library(ggplot2)

ggplot(startup_funding, aes(x = Sector, y = Funding)) +
  geom_boxplot(fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Funding Distribution Across Sectors")
```

---

### Revenue Distribution

```{r}
ggplot(startup_funding, aes(x = Revenue)) +
  geom_histogram(bins = 30, fill = "orange") +
  ggtitle("Revenue Distribution")
```

---

### Correlation Analysis

```{r}
library(GGally)
library(dplyr)

num_data <- startup_funding %>%
  select(Funding, Revenue, Employee, Marketing, Operational_Years)

GGally::ggpairs(num_data)
```

---

## 5. Multivariate Normality Test

```{r}
library(MVN)

mvn_res <- startup_funding %>%
  select(Funding, Revenue, Marketing)

mvn_dataRes <- mvn(data = mvn_res, mvn_test = "mardia")

mvn_dataRes$multivariate_normality
```

**Interpretation:**
The skewness test indicates deviation from multivariate normality, while kurtosis remains acceptable.
This suggests real-world skewness in financial data.

---

## 6. Principal Component Analysis (PCA)

```{r}
pca_data <- scale(startup_funding %>%
                   select(Funding, Revenue, Employee, Marketing))

pca <- prcomp(pca_data, center = TRUE, scale. = TRUE)

summary(pca)
```

---

### Scree Plot

```{r}
library(factoextra)
fviz_eig(pca)
```

---

### PCA Loadings

```{r}
pca$rotation
```

---

### Extract Components

```{r}
startup_funding$Company_Scale <- pca$x[,1]
startup_funding$Operational_Intensity <- pca$x[,2]
```

**Interpretation:**

* PC1 represents **Company Scale**
* PC2 represents **Operational Intensity**

---

## 7. ANOVA

```{r}
anova_model <- aov(Company_Scale ~ Sector, data = startup_funding)
summary(anova_model)
```

---

### Post Hoc Test

```{r}
TukeyHSD(anova_model)
```

---

### Visualization

```{r}
ggplot(startup_funding, aes(x = Sector, y = Company_Scale)) +
  geom_boxplot(fill = "lightgreen") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

## 8. Conclusion

* Startup scale does not significantly differ across sectors (p > 0.05)
* Growth patterns are not strongly dependent on industry
* PCA effectively reduced multiple variables into meaningful components
* Startup success is multi-dimensional and influenced by multiple factors

---

## 9. Save Processed Data

```{r}
write.csv(startup_funding, "processed_startup_data.csv", row.names = FALSE)
```
